Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-03-02 supplement.tex
Fix remarks to optimized complexity

Remarkably, this strongly improves the algorithm's time
complexity. Where the original algorithm enumerates all partial
sequences $\substitute{\val}{\val'}$, where
$\val\in\partseqs(\separator{u}{v})$ and
$\val'\in\partseqs(\difference{u}{v})$, the modified algorithm
enumerates in two stages:
\begin{enumerate}
\item for each connected component of the base pair dependency graph
  that involves any position in
  $\separator{u}{v}\cup\difference{u}{v} = \chi(v)$, it first
  enumerates all possibilities how the positions from the partition
  sets are assigned to either $\{\Ab,\Gb\}$ or $\{\Cb,\Ub\}$
\item it enumerates the possible partial sequences from the (according
  to the decisions of the first stage) restricted domains of the
  position in $\chi(v)$.
\end{enumerate}
Let $c(v)$ be the number of connected components involving positions
in $\chi(v)$; then, the first stage enumerates $2^{c(v)}$
combinations. For each such combination, the second stage enumerates
only $2^{|\chi(v)|}$ partial sequences, since there remain only two
possible nucleotide choices for each position. All message values for
node $v$ are thus computed in time
$\Theta(2^{c(v)+|\chi(v)|})$. Consequently, the total time complexity
of the modified algorithm is $$O(nk2^{\hat w}),$$ where $n$ is the RNA
length, $k$ is the number of structures, and
$\hat w := \max_{v\in T} c(v)+|\chi(v)|$. Notably, $\hat w$ can not
exceed $2(w+1)$ such that the modified algorithm does not increase the
original complexity.
%
As useful corollary, one obtains the (looser) complexity
bound $$O(nk2^{w+1}2^{cc})$$ in terms of treewidth $w$ and connected
components $cc$.

>>>>

Remarkably, this strongly improves the algorithm's time
complexity. Where the original algorithm enumerates all partial
sequences $\substitute{\val}{\val'}$, where
$\val\in\partseqs(\separator{u}{v})$ and
$\val'\in\partseqs(\difference{u}{v})$, the modified algorithm
enumerates in two stages:
\begin{enumerate}
\item for each connected component of the base pair dependency graph
  that involves any position in
  $\separator{u}{v}\cup\difference{u}{v} = \chi(v)$, it first
  enumerates all possibilities how the positions from the partition
  sets are assigned to either $\{\Ab,\Gb\}$ or $\{\Cb,\Ub\}$
\item it enumerates the possible partial sequences from the (according
  to the decisions of the first stage) restricted domains of the
  position in $\chi(v)$.
\end{enumerate}
Let $c(v)$ be the number of connected components involving positions
in $\chi(v)$; then, the first stage enumerates $2^{c(v)}$
combinations. For each such combination, the second stage enumerates
only $2^{|\chi(v)|}$ partial sequences, since there remain only two
possible nucleotide choices for each position. All message values for
node $v$ are thus computed in time
$\Theta(2^{c(v)+|\chi(v)|})$. Consequently, the total time complexity
of the modified algorithm is $$O(nk2^{\hat w}),$$ where $n$ is the RNA
length, $k$ is the number of structures, and
$\hat w := \max_{v\in T} c(v)+|\chi(v)|$.
%
From this directly follows the slightly looser complexity
bound $$O(nk2^{w+1}2^{c})$$ in terms of treewidth $w$ and the maximum
number $c$ of connected components represented in a node of the tree
decomposition.  Notably, $\hat w$, as well as $w+1+c$, can not exceed
$2(w+1)$, since $c\leq w+1$. Consequently the modified algorithm does
never increase the original complexity.
